A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization

### 2013-09-04 reviewer_response.tex

\rev{This idea of PSG was mentioned but was not covered in 
the paper. For completeness and value proposition sake, I would recommend to 
add at least half section to show how it will help and look like. You can 
take the simplest formulation of your choice and show how MCG and FPG will 
translate and look as a PSG. A brief discussion of advantages vs 
disadvantages will be useful. }

We agree that this was too quickly addressed in the paper. While we feel that 
discussion of the PSG in any significant depth is too big of a topic to fit into 
this paper, it does warrant a more complete description than was given in the 
original draft.

We have added sections 3.5, 4.4 in the appropriate 
parts of the paper to address the the driver nodes and driven edges which 
together distinguish a PSG from and FPG.  We have also added added section 5.5
which gives an example of a simple PSG and compares a PSG with its close relative
and XDSM diagram.

It is our conclusion that the PSG represents a sub-set of the information in the XDSM
diagram, but that for data connectivity part of the graphs XDSM and PSG represent 
equivalent graphs. It would be interesting future work to consider incorporating a new
process edge type into the PSG to make it functionally equivalent to a complete XDSM, 
however that would be a significant amount of extra work and worth of it's own paper. 
To do that topic justice, the original XDSM work would need to be converted for every 
case considered into the proposed syntax to ensure that they were completely equivalent.

@@

\rev{ The idea of “holes” and “collision” is explained a few times in the paper. 
It can be reduced. }

Searching the paper, the concepts is discussed often, but only formally defined once 
in section 4.1. In 5.3 a mathematical description of the set of nodes that meet
the criteria for holes and edges is given, but this is different than the definition.

We have checked both sections and removed the gross reducency, but feel that keeping 
both equations makes for a clearer message in the paper. The concept is a bit subtle 
since a node can be either a design variable or a hole, but not both. Similarlly, 
a collision can occur when a variable has two incomming edges or not if the problem 
is multi-fidelity. So a little extra explanation helps to keep this clear.

@@

\rev{The review of other graph based techniques is good and highlights their 
limitations, but I would like to see a discussion of what challenges a 
practitioner may face while trying to use the approach suggested in this 
paper. The MCG and FPG will cover lot of information like objectives and 
constraints, but at what expense. Comparing XDSM with MCG or FPG, one might 
find XDSM as more easy to read and comprehend. For a similar size problem, 
FPG looks far more complex and bigger.}

We agree that XDSM is more visually informative and consise. However this was not the 
goal of the new syntax. Instead, the goal was to make a graph structure that 
was easier to operate on algorithmically. As a consequence the graph grows larger and 
less human readable. This trade is acceptable for two reasons:

\begin{enumerate}
    \item It should be possible to convert between the XDSM format and the proposed syntax 
    in a lossless manner. This is true for the gray lines of XDSM but not the black ones.
    Only data connections are shown in current syntax. 
    \item Algorithmic methods are a useful way of analyzing problem structure and it makes 
    sense to have a graph structure that works well for this use, even at the expense 
    of human readability. 
\end{enumerate}

See the revised end of the introduction (section 1) where this is more clearly stated.

### 2013-09-04 graph_syntax.tex
more revisions

  Driven edges are the distinguishing characteristic associated with driver nodes and
  driver blocks. Driven edges don't follow the same rules as other edges. Driven edges  
  will always connect a driver node to a variable node (in either direction), but they 
  are not considered when counting the indegree or outdegree of the variable node.
  Hence a single variable node could have multiple incoming driven edges from different 
  driver blocks. Figure \ref{f:driver block} shows a notional example where a driver block 
  represents and optimizer solving an unconstrained minimization problem.

>>>>

  Driven edges are the distinguishing characteristic associated with driver nodes and
  driver blocks. Driven edges don't follow the same rules as other edges. Driven edges  
  will always connect a driver node to a variable node (in either direction), but they 
  are not considered when counting the indegree or outdegree of the variable node.
  Hence a single variable node could have multiple incoming driven edges from different 
  driver blocks, e.g. in a sequential optimization strategy, both drivers would at 
  different times, pick values for the same variable. Figure \ref{f:driver block} shows 
  a notional example where a driver block represents and optimizer solving an 
  unconstrained minimization problem.

### 2013-09-04 intro.tex
more revisions

    With the added complexity from larger models, it is plausible that the task of combining all the analyses into a 
    consistent system model capable of solving a relevant engineering design 
    problem could approach the cost and time requirements of creating any of the discipline 
    analyses themselves. For these large scale MDO problems, the couplings between the 
    analyses begin to dominate the effort required in setting up the model. It is this problem of 
    determining sets of analysis tools and their interconnectivities to form realistic
    multidisciplinary problems that is the subject of this paper. We are motivated by the following notional
    but realistic problem of organizing an MDO study for a complex system:

@@

    This information in an FPG represents the engineering design problem that needs to be solved, but it is not 
    itself sufficient to actually run a design optimization. It does not provide infomration 
    about how to solve the problem. Even after the problem is defined an integration platform or framework needs 
    to be selected and appropriate optimization methods identified. This last step essentially 
    involves taking the FPG and turning it into a usable model. We breifly consider methods to further 
    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be usefull in this 
    task, but that is not the focus of this paper. The work is primarily concerned with applying graph 
    theory to the creation of an FPG and the benefits to the design process by doing so.

@@

\subsection{Background in Graph-Based Descriptions of MDAO Problems}
	\label{s:existing syntax}
    As indicated in the examples in Sec.~\ref{s:specific vs fundamental}, the mathematical 
    language for specifying optimization problem formulations is very general and can be used both for 
    fundamental and specific problem formulations. Tedford and Martins used this syntax to specify the 
    FPF for a set of test problems and also to describe specific formulations for solving them with a 
    number of optimization architectures \cite{Tedford2009}. By solving different specific problem statements corresponding to the same FPF, they were able to  benchmark the performance of different optimization architectures against a fixed set of 
    problems. Gray et al. also benchmarked the performance of a set of MDAO architectures  with a similar approach \cite{Gray2013}. 
    This work examined a larger set of architectures and proposed the use of OpenMDAO as a platform to build a larger community-developed set of test problems and architectures. Both of these benchmarking efforts 
    demonstrate how multiple specific problem formulations can relate to a common FPF and indicate the 
    value of a common problem descrition. The challenge with this 
    traditional syntax is that it is not easily manipulated or analyzed with automatic procedures to explore alternate problem formulations.

@@

    Alexandrov and Lewis introduced a graph based syntax called REMS which 
    incorporates objectives and constraints into a graph, effectively combining 
    FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
    matrix form from DSM, but by adding the objectives and constraints, it partially 
    combines a traditional DSM with an FDT. This formulation allows REMS to represent data 
    dependency between multiple analysis tools as well as between analysis tools and
    objective/constraint functions. Additionaly, REMS address the need to
    transition between multiple solution strategies while maintaining a single consistent  
    graph representation of the fundamental problem formulation. REMS provides
    syntax for variables and functions but does not facilitate inclusion 
    of solvers or optimizers in the graph.  The 
     graph syntax proposed in this paper could  
    be viewed as an extention to REMS with the goal of allowing the graphs to  describe analysis couplings and more 
    closely interact with existing design processes and MDAO frameworks.

@@

In this paper, we propose a new graph syntax that 
    combines certain features of REMS and XDSM. 
The approach allows interactions and couplings of analysis tools and objective and constraint functions to be described while retaining the flexiblity to 
    represent a fundamental problem formulation free of solution specific information.

\subsection{Requirements for a New Graph Syntax}
  \label{s:requirements}
  The goal of the graph-based syntax presented here is to enable the general 
  structure of an MDAO problem to be described independently of any solution information, 
  while still being able to accomodate the more specific case when a solution 
  strategy is applied. In order to achieve this goal, 
  the graph syntax needs to accomodate a number of MDAO problem constructs: 
  \begin{itemize}
    \item Analysis tools and their interconnections
    \item Design variables, objectives, and constraints
    \item Coupling between analyses
    \item Multi-fidelity analyses
  \end{itemize}

>>>>

    With the added complexity from larger models, it is plausible that the task of combining all the analyses into a 
    consistent system model capable of solving a relevant engineering design 
    problem could approach the cost and time requirements of creating any of the discipline 
    analyses themselves. For these large scale MDO problems, the couplings between the 
    analyses begin to dominate the effort required in setting up the model. It is this problem of 
    determining sets of analysis tools and their inter-connectivities to form realistic
    multidisciplinary problems that is the subject of this paper. We are motivated by the following notional
    but realistic problem of organizing an MDO study for a complex system:

@@

    This information in an FPG represents the engineering design problem that needs to be solved, but it is not 
    itself sufficient to actually run a design optimization. It does not provide information 
    about how to solve the problem. Even after the problem is defined an integration platform or framework needs 
    to be selected and appropriate optimization methods identified. This last step essentially 
    involves taking the FPG and turning it into a usable model. We briefly consider methods to further 
    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be useful in this 
    task, but that is not the focus of this paper. The work is primarily concerned with applying graph 
    theory to the creation of an FPG and the benefits to the design process by doing so.

@@

\subsection{Background in Graph-Based Descriptions of MDAO Problems}
	\label{s:existing syntax}
    As indicated in the examples in Sec.~\ref{s:specific vs fundamental}, the mathematical 
    language for specifying optimization problem formulations is very general and can be used both for 
    fundamental and specific problem formulations. Tedford and Martins used this syntax to specify the 
    FPF for a set of test problems and also to describe specific formulations for solving them with a 
    number of optimization architectures \cite{Tedford2009}. By solving different specific problem statements corresponding to the same FPF, they were able to  benchmark the performance of different optimization architectures against a fixed set of 
    problems. Gray et al. also benchmarked the performance of a set of MDAO architectures  with a similar approach \cite{Gray2013}. 
    This work examined a larger set of architectures and proposed the use of OpenMDAO as a platform to build a larger community-developed set of test problems and architectures. Both of these benchmarking efforts 
    demonstrate how multiple specific problem formulations can relate to a common FPF and indicate the 
    value of a common problem description. The challenge with this 
    traditional syntax is that it is not easily manipulated or analyzed with automatic procedures to explore alternate problem formulations.

@@

    Alexandrov and Lewis introduced a graph based syntax called REMS which 
    incorporates objectives and constraints into a graph, effectively combining 
    FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
    matrix form from DSM, but by adding the objectives and constraints, it partially 
    combines a traditional DSM with an FDT. This formulation allows REMS to represent data 
    dependency between multiple analysis tools as well as between analysis tools and
    objective/constraint functions. Additionally, REMS address the need to
    transition between multiple solution strategies while maintaining a single consistent  
    graph representation of the fundamental problem formulation. REMS provides
    syntax for variables and functions but does not facilitate inclusion 
    of solvers or optimizers in the graph.  The graph syntax proposed in this paper could  
    be viewed as an extension to REMS with the goal of allowing the graphs to  describe analysis couplings and more 
    closely interact with existing design processes and MDAO frameworks.

@@

    In this paper, we propose a new graph syntax that combines certain features of REMS and XDSM. 
    Despite sharing some features with both of these other graph based approaches, the syntax
    proposed here has a fundamentally different goal which is complimentary to both of them. 
    REMS and XDSM provide excellent human readable formats for an MDAO problem description. This 
    syntax is designed to provide a more effective machine readable format for MDAO problem definition. 
    In the interest of making it easier to algorithmically operate on the graphs, the specified 
    format is both more verbose and less visually informative. The benefit is a more consistent 
    and easily utilized graph from a algorithm perspective.

\subsection{Requirements for a New Graph Syntax}
  \label{s:requirements}
  The goal of the graph-based syntax presented here is to enable the general 
  structure of an MDAO problem to be described independently of any solution information, 
  while still being able to accommodate the more specific case when a solution 
  strategy is applied. In order to achieve this goal, 
  the graph syntax needs to accommodate a number of MDAO problem constructs: 
  \begin{itemize}
    \item Analysis tools and their interconnections
    \item Design variables, objectives, and constraints
    \item Coupling between analyses
    \item Multi-fidelity analyses
  \end{itemize}

### 2013-09-04 mcg_fpg.tex
more revisions

\begin{description}
    \item[\bf{Step 1: Holes}] 
        The first step is to detect holes and disconnect the first set of connection edges downstream of them, as indicated in Fig.~\ref{f:hole}. 
        These connection edges are removed because they represent output variables which cannot be determined because the analysis function does not have adequate inputs.

@@

\subsection{Problem Solution Graph}
Once a suitable FPG has been obtained the issue of how to structure a solution strategy 
can be addressed. Similar to the way an MCG can be manipulated into a FPG, an FPG can be
further manipulated into a PSG. To get from an MCG to an FPG you eliminated analysis blocks and
connection edges to reduce the problem to something solvable, but to get from an FPG to a PSG you 
add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
to the problem.

>>>>

\begin{description}
    \item[\bf{Step 1: Holes}] 
    The first step is to detect holes and disconnect the first set of connection edges 
    downstream of them, as indicated in Fig.~\ref{f:hole}.

@@

\subsection{Problem Solution Graph}
Once a suitable FPG has been obtained the issue of how to structure a solution strategy 
can be addressed. Similar to the way an MCG can be manipulated into a FPG, an FPG can be
further manipulated into a PSG. To get from an MCG to an FPG you eliminated analysis blocks and
connection edges to reduce the problem to something solvable, but to get from an FPG to a PSG you 
add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
to the problem. For example, Figure \ref{f:selllar_psg} shows the PSG that solves the Sellar problem 
with an MDF solution strategy. There are two driver blocks in this PSG, one for the solver that 
ensures compatibility and another for the optimizer that picks the design variables.

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=3.0in]{images/sellar_psg}
    \end{center}
    \vspace{-10pt}
    \caption{Example PSG for solving the Sellar problem with MDF strategy}
    \label{f:sellar_psg}
\end{figure}

The PSG in figure \ref{f:selllar_psg} can be compared to the XDSM for the same situation 
in Fig. \ref{fig:XDSM_simple}. While the two diagrams contain much of the same information 
there is a significant difference. The PSG describes the data passing that happens between 
the drivers and the rest of the problem, but it does not describe when that data passing 
will occur (the execution order of the blocks). In the proposed graph syntax edges describe 
the passing of data between one node an another, but they do not define when that 
communication happens.

XDSM diagrams do contain information about data passing as well as when that data should be passed. 
XDSM essentially represents two related, but independent graphs simultaneously. 
This is accomplished by including two distinct sets of edges, one for data 
connections and a second indicating execution sequence. As seen in Fig. \ref{fig:XDSM_simple}
data connections are bold-gray edges and execution connections are numbered-black edges. The
execution edges, the back edges, in XDSM represent and additional graph that
follows a completely different set of rules and behaviors from the data connections graph. 
On the other hand the data described by the gray edges, 
the data passing information, is equivalent to the information in the PSG. XDSM presents this 
information in a more compact and human readable manner by aggregating groups of related variables
into combined nodes. The more verbose form presented in a PSG is less human readable, but more easily 
machine readable as a consequence of a more consistent structure.
